---
title: "Bootstrap"
output: html_notebook
---

#Bootstrapping Approach

Bootstrapping is a resampling technique to estimate statistics on a population by sampling a dataset with replacement.

A useful feature of the bootstrap method is that the resulting sample of estimations often forms a Gaussian distribution. In additional to summarizing this distribution with a central tendency, measures of variance can be given, such as standard deviation and standard error. Further, a confidence interval can be calculated and used to bound the presented estimate. This is useful when presenting the estimated skill of a machine learning model. (https://machinelearningmastery.com/a-gentle-introduction-to-the-bootstrap-method/)

Seth has proposed that the best way of using this approach is to sample the genes at random, n being the number of canon or noncanon genes that were fed into that run. The count will then be the number of genes under selection that occur in that sample. These genes being the genes - regardless of class - that were found to be under selection in that lineage codeml run. Though I'm not sure if I wholly understand the logic yet, the proposal is that this will produce a distribution of "background selection" against which we can there compare the number of canon / noncanon genes under selection against. 

I'm going to test this out on an origin that has fewer genes under selection, one that has many, and then one somewhere in between. 

First, I need a table that tells me all the descriptive stats (such as genes under selection per lineage), so lets start there.

##Getting Counts

```{r}
summary(data.all)
```

Check for duplicates

```{r}
nrow(data.all) == nrow(data.all[!duplicated(data.all),])
```

Test this out.

Pick three origins at random

```{r}
soc.list <- data.all$SocOrigin[!duplicated(data.all$SocOrigin)]

test1 <- data.all[data.all$SocOrigin == paste(soc.list[2]),]
test2 <- data.all[data.all$SocOrigin == paste(soc.list[8]),]
test3 <- data.all[data.all$SocOrigin == paste(soc.list[5]),]
```

Check specifically there are no gene duplications within each

```{r}
nrow(test1) == nrow(test1[!duplicated(test1$Gene),])
nrow(test2) == nrow(test2[!duplicated(test2$Gene),])
nrow(test3) == nrow(test3[!duplicated(test3$Gene),])
```


Ok I trust it. Moving on.

Let's write a function that will go through each of the tested lineages, extract them from the master dataframe and count genes in, genes under selection per class, etc.

```{r}
col.headers <- c("LineageTested", "GenesInput", "CanonInput", "NonCanonInput", "RandomInput", "GenesUnderSel", "CanonUnderSel", "NonCanonUnderSel", "RandomUnderSel")

out <- matrix(ncol = length(col.headers), nrow = length(soc.list))


for (i in 1:length(soc.list)){
  ext <- data.all[data.all$SocOrigin == paste(soc.list[i]),]
  soc <- paste(soc.list[i])
  total <- nrow(ext)
  tot.can <- as.numeric(nrow(ext[ext$Class == "Canon",]))
  tot.non <- nrow(ext[ext$Class == "NonCanon",])
  tot.ran <- nrow(ext[ext$Class == "Random",])
  tot.sel <- nrow(ext[ext$adj_pvalue < 0.05,])
  sel.can <- nrow(ext[ext$Class == "Canon" & ext$adj_pvalue < 0.05,])
  sel.non <- nrow(ext[ext$Class == "NonCanon" & ext$adj_pvalue < 0.05,])
  sel.ran <- nrow(ext[ext$Class == "Random" & ext$adj_pvalue < 0.05,])
  chain <- c(soc, total, tot.can, tot.non, tot.ran, tot.sel, sel.can, sel.non, sel.ran)
  out[i,] <- chain
}

master.count <- as.data.frame(out)
colnames(master.count) <- col.headers
master.count
```


Now for the bootstrapping

##Bootstrap

###Test

I'm going to write this off-script, and test it out on a lineage with lots of selection first.

```{r}
master.count[which.max(master.count$GenesUnderSel),]
```

```{r}
rownames(master.count) <- master.count$LineageTested
master.count
```


For the canon genes I need to:

- make a list of all genes that were input into the Meli run

- specify the number of canon genes that were input which will be used as the "n" of genes to be sampled from the background list of genes

- make a list of all the genes that were considered under selection in the Meli run

- take note of the number of genes picked at random that were under selection in the Meli run

- put the sample back into the "population" and sample again (x1000 to test)

Canon 

```{r}
dummy.sel <- function(x,y){
  class <- x
  soc <- y
  if ( class == "Canon" || class == "NonCanon"){
    des <- "CanonInput"
  } else {
    des <- "NonCanonInput"
  }
  pop <- droplevels(data.all$Gene[data.all$SocOrigin == paste(soc)])
  n <- master.count[soc,des]
  game <- droplevels(data.all$Gene[data.all$SocOrigin == paste(soc) &
                                     data.all$adj_pvalue < 0.05])
  catch <- sample(pop, size=n, replace = T)
  hit <- sum(as.numeric(game %in% catch))
  return(hit)
}
  
meli.out <- rep(0,999)
for (i in 1:999){
  meli.out[i] <- dummy.sel("Canon", "Meli")
}

hist(meli.out)
```

So it works making distributions with lineages that have a relatively large number of genes exhibiting signs of positive selection ... but is it going to work when there are small numbers of genes under selection 

(I'm guessing not)

```{r}
master.count[which.min(master.count$GenesUnderSel),]
```

```{r}
corb.epi.out <- rep(0,999)
for (i in 1:999){
  corb.epi.out[i] <- dummy.sel("Canon", "CorbSocEpi")
}

hist(corb.epi.out)
```

I. Was. Right. Unfortunately.

```{r}
49*.85
```

