---
title: "GO:Analysis"
output: html_notebook
---

I'm going to start by makeing tables of genes of interest - meaning all genes that have appeared as under selection in more than one lineage. I will do this separately for solitary and social lineages before combining at the end. 

GO analysis will be ran on canon / noncanon / random genes under selection per lineage. Go terms can then be checked for comparisons between socialities, etc.

##Genes of Interest

What genes are worth looking at ? Are there genes that pop up in more than one origin / elaboration of sociality, suggesting the change in function may have been selected for in accordance with transitioning to more social lifestyles ? And if so, what are these genes?

This script will take the data, keep only those significant and aggregate by SocOrigin in an object called "undersel". All of those genes that appear more than once will be put in a more interesting dataframe called "goi"

The rest will be put in a dataframe called "undersel"

```{r}
source("Scripts/getgoi.R", local = knitr::knit_global())
```

```{r}
head(goi)
```

Should probably remove the random genes.

```{r}
goi.imm <- goi[!goi$Class == "Random",]
```

```{r}
head(goi.imm)
```

```{r}
goi.imm[goi.imm$Class == "Canon",]
```

```{r}
goi.imm[goi.imm$count > 2,]
```

```{r}
nrow(goi.imm[goi.imm$count > 1,])
nrow(goi.imm[(goi.imm$count > 1) & (goi.imm$Class == "Canon"),])
nrow(goi.imm[(goi.imm$count > 1) & (goi.imm$Class == "NonCanon"),])
```

```{r}
write.table(goi, "output/GOI.tsv", sep = "\t", row.names = F, quote = F)
```

And to do the same for solitary ...

```{r}
source("Scripts/getgoiSol.R", local = knitr::knit_global())
```

```{r}
head(goi.sol)
```

Store the immune genes separately and then write up the solitary GOI.

```{r}
goi.sol.imm <- goi[!goi$Class == "Random",]

write.table(goi, "output/GOI_Sol.tsv", sep = "\t", row.names = F, quote = F)
```


Combine the dataframes and re-aggregate
Remove count and recalculate

```{r}
goi.all <- rbind(goi, goi.sol)

goi.all <- aggregate(goi.all[2], goi.all[-2],
                     FUN = function(X) paste(unique(X), collapse=", "))

goi.all$count <- NULL

goi.all$count <- str_count(goi.all$SocOrigin, pattern = ",")+1

head(goi.all)
```

What's the most often to experience positive selection on immune - related genes? 

```{r}
goi.all[!goi.all$Class == "Random" & goi.all$count > 3,]
```

Write up

```{r}
write.table(goi.all, "output/GOI_All.tsv", sep = "\t", col.names = T, row.names = F, quote = F)
```


##Canon Function

Looking at the canon genes that are under selection (above or equal to one lineage at a time) and assign immune function.

In the previous GOI scripts genes under selection were saved in objects undersel and undersel.sol respectively. 

```{r}
undersel.all <- rbind(undersel, undersel.sol)
undersel.all <- aggregate(undersel.all[2], undersel.all[-2],
                          FUN = function(X) paste(unique(X), collapse = ", "))
undersel.all$count <- NULL
undersel.all$count <- str_count(undersel.all$SocOrigin, pattern = ",")+1
tail(undersel.all)
```

Out of all the genes assessed, 1950 genes total were found under positive selection.

```{r}

```



Assign function to the canonical immune genes and save

```{r}
imm.class <- read.table("ImmResources/Canon_Immune_GeneList_Jan20.txt", header = T, sep="\t", quote = "")
imm.fun <- imm.class[,c(1:2)]
head(imm.fun)
```


```{r}

```

