---
title: "UpSetR_Workbook"
output: html_notebook
---

#Comparisons

Now that I have GO terms and genes under selection, its prudent to compare. Let's see what upsetR is saying.

```{r}
install.packages("UpSetR")
library("UpSetR")
```

I've written a script that does everything I need here.

1. Makes tables of GOterms based on absence / presence (1/0) in each lineage/sociality

2. Makes master tables and separate immune / random only tables

3. Write these tables in an output folder 

```{r}
source("Scripts/makeUpsetR.R", local = knitr::knit_global())
```

All Upset plots were made in the Script MakePlots.R

I have new GO term lists from ReViGo which have reduced the GO terms to more broad parent terms. It will be using this I'll be making any manuscript level figures with but I need to make the tables I needed for UpSetR to do so.

I manually picked out terms that looked to be immune from the ReViGo results and put them into a list (below).

```{r}
go.immune <- read.table("Genome_Misc/ImmGOterms_GUS.txt", header = F)
head(go.immune)
```

```{r}

```


